% !TEX root = ln_diss.tex
\chapter*{Acknowledgements}

During these years of Ph.D. studies several persons (and I may forget many of them) have contributed to my work and without their support it would have been impossible for me to complete it. That is why I wish to dedicate this section to recognize their support.

I would like to express my sincere gratitude to the professors of my advisors: Prof. Giuseppe Pelosi and Dr. Stefano Selleri, for their guidance, motivation and support during this period.

I would like to acknowledge the financial support from Nuovo Pignone -- GE Oil and Gas, the research here presented being a part of my studies on electromagnetic compatibility modeling techniques.

Special thanks to all the employees, researchers and students of the Computational Electromagnetics laboratory of the Information Engineering department at the University of Florence, for their courtesy, the pieces of advice and help they gave me, and most of all for the good time spent together.

I would also like to express my appreciation to the pieces of knowledge also transferred to me all along my finite element methods for microwave engineering mastering quest by Prof. Romanus Dyczij-Edlinger, Prof. Jin Fa-Lee and Prof. Oszk\'ar B\'ir\'o. Deep appreciation also to Dr. Ortwin Farle, for many implementation skills imparted to me during my master thesis work at the \textit{Lehrstuhl f\"ur Theoretische Elektrotechnik} laboratory of Saarland University.

I would also like to express my heartfelt gratitude to my family for encouragement
and continuous support given throughout my life.

Last but not least, I would like to express my appreciation to my sweetheart, Veronica, for her love, patience, support and understanding throughout these difficult and challenging years.