% !TEX root = ln_diss.tex
\chapter*{List of Publications}%\label{publications}
%\markboth{List of Publications}{}
%\addcontentsline{toc}{chapter}{List of Publications}
%\label{page:publications}
\section*{Refereed papers}
%\addcontentsline{toc}{section}{Refereed Journal Papers}
\begin{enumerate}[label={[\arabic*]}]
\item  L. Ntibarikure, G. Pelosi, S. Selleri, ``Efficient harmonic balance analysis of waveguide devices with nonlinear dielectrics,'' \emph{Microwave and Wireless Components Letters, IEEE}, vol. 22, no. 5, pp. 221--223, 2012.
\item  L. Ntibarikure, G. Pelosi, S. Selleri, ``Harmonic balance domain decomposition finite elements for nonlinear passive microwave devices analysis,'' \emph{Special Issue on \quotes{Finite Elements for Microwave Engineering}, Electromagnetics}, vol.~34, no. 3, 2014.
\end{enumerate}

\section*{Conference proceedings}
%\addcontentsline{toc}{section}{Refereed Journal Papers}
\begin{enumerate}[label={[\arabic*]}]
\item  L. Ntibarikure, ``Model order reduction in finite element analysis of phased array antennas,'' in \emph{Numero Speciale 8 - Serie di Elettromagnetismo su \quotes{XIX Riunione Nazionale di Elettromagnetismo (RiNEm)}, Atti della \quotes{Fondazione Giorgio Ronchi}}, vol. LXVIII, no. 4, pp. 97--102, 2012. [Barzilai award 2012]
\end{enumerate}


\section*{Conferences}
%\addcontentsline{toc}{section}{Refereed Conference Papers}
\begin{enumerate}[label={[\arabic*]}]
\item L. Ntibarikure, G. Pelosi, and S. Selleri, ``Harmonic balance domain decomposition finite element for nonlinear passive microwave devices analysis,'' in \emph{11th International Workshop on Finite Elements for Microwave Engineering -- FEM2012}, Estes Park, Colorado, USA, Jun. 4--6, 2012.
\item L. Ntibarikure, G. Pelosi, S. Selleri, ``Harmonic balance finite element analysis of third order intermodulation products in ferrite devices,'' in \emph{XIX Riunione Nazionale di Elettromagnetismo -- XIX~RiNEm}, Rome, Italy, Sep. 10--14, 2012.
\item  L. Ntibarikure, ``Model order reduction in finite element analysis of phased array antennas,'' in \emph{XIX Riunione Nazionale di Elettromagnetismo -- XIX~RiNEm}, Rome, Italy, Sep. 10-14, 2012.
\end{enumerate}
